---
title: "Transform TIDAL results to planar C:P"
author: "irw"
date: "Created 20240207 -- Last compiled `r format(Sys.Date())`"
always_allow_html: false
output:
  html_document:
    df_print: paged
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
    keep_tex: false
params:
  verbose: yes
changes:
- "Adapted from report template"
editor_options: 
  chunk_output_type: inline
---

```{r, include = F}

child_chunk = knitr::opts_knit$get("child")

```


```{r, eval = !child_chunk, include=FALSE}
if (params$verbose) {
  knitr::opts_chunk$set(
    echo = TRUE,
    warning = TRUE,
    message = FALSE
  )
} else {
  knitr::opts_chunk$set(
    echo = FALSE,
    warning = FALSE,
    message = FALSE
  )
  }



knitr::opts_chunk$set(fig.width=120*1/25.4, fig.height=90*1/25.4, 
                      fig.retina = 1, dpi=500)
```


```{r, echo=FALSE}
if (knitr::is_latex_output()) {
knitr::asis_output("\\newpage")
}
```

```{r, eval = !child_chunk}
# Setup chunk

# Load libraries
library(tidyverse)
library(here)
library(jsonlite)
library(kableExtra)
library(plotly)
library(glue)
library(pracma)
library(car)

source(here::here("resources/utils/make_path.R"))
source(here::here("resources/utils/capitalize.R"))
source(here::here("resources/plot/print_ggplot.R"))
source(here::here("resources/plot/plot_interactive.R"))

knitr::asis_output("# Session Info")
sessionInfo()
```

```{r, eval = !child_chunk, echo=FALSE}
if (knitr::is_latex_output()) {
knitr::asis_output("\\newpage")
}
```

```{r}
parent <- "deposition/cp-transform"
```

Original reference:

-	Schroeter JD, Pritchard JN, Hwang D, Martonen TB. Airway identification within planar gamma camera images using computer models of lung morphology. Pharm Res. 2005;22(10):1692-9.

NOTE: This is based on a symmetric airway model, using Weibel numbering.

Fraction reference:

- On the Validation of Generational Lung Deposition Computer Models Using Planar Scintigraphic Images: The Case of Mimetikos Preludium. Bo Olsson and Stavros C. Kassinos. Journal of Aerosol Medicine and Pulmonary Drug Delivery 2021 34:2, 115-123


```{r}
cip_frac <- read_csv(make_path("res/map.csv"))

cip_transform <- cip_frac %>% select(-G) %>% t()

# Split I equally between c, p
cp_frac <- cip_frac %>% 
  mutate(
    c = C+I/2,
    p = P+I/2,
    s = c+p
  )

cp_transform <- cp_frac %>% 
  select(c,p) %>% 
  t()
  
```

Average generation regions used in TIDAL algorithm

C: G3-G8
P: G9+

```{r}
# Convention: trueRegion_tidalRegion

g_start = 3
g_end = 8

# Get the fraction of regional deposition from the c region in TIDAL
cp_frac_ctidal <- cp_frac %>% 
  slice(g_start:g_end) %>% 
  select(c,p) %>% 
  summarize(across(.cols = c(c,p), mean, .names = "{.col}"))

# Get the fraction of regional deposition from the p region in TIDAL
cp_frac_ptidal <- cp_frac %>% 
  slice(g_end+1:n()) %>% 
  select(c,p) %>% 
  summarize(across(.cols = c(c,p), mean, .names = "{.col}"))

cp_transform <- bind_rows(
  cp_frac_ctidal,
  cp_frac_ptidal
  ) %>% 
  t() 

# Confirm that fractions add to 1
cp_transform[,1] %>% sum()
cp_transform[,2] %>% sum()

# Check condition number
cp_transform %>% kappa()

```


```{r}
cp_tidal <- 0.8
p_basis <- 1

c_tidal <- cp_tidal * p_basis
p_tidal <- p_basis

# Planar C:P
cp_transform %*% matrix(c(c_tidal, p_tidal)) %>% 
  (function(x) {x[1]/x[2]})
```

```{r}
# Make it a function
convert_to_planar <- function(cp_tidal, p_basis = 1) {
  c_tidal <- cp_tidal * p_basis
  p_tidal <- p_basis
  
  cp_planar <- cp_transform %*% matrix(c(c_tidal, p_tidal)) %>% 
  (function(x) {x[1]/x[2]})
  
  return(cp_planar)
}

cp_tidal_series = seq(0,10,0.1);
plot(cp_tidal_series, lapply(cp_tidal_series, convert_to_planar))
```

# Compare experimental results

```{r}
exp_data <- tibble(
  designator = c("shallow", "deep", "uniform"),
  tidal = c(0.69, 0.53, 1.75)
) %>% mutate(
  planar = map(tidal, convert_to_planar) %>% unlist()
)

exp_data
```

## Plot

```{r, fig.width = 170/3/25.4, fig.height = 40/25.4}
knitr::opts_chunk$set(fig.width = 170/3/25.4, fig.height = 40/25.4)

conversion_plot <- (
  tibble(
  x = cp_tidal_series,
  y = lapply(x, convert_to_planar) %>% unlist()
) %>% 
  ggplot() +
  geom_line(aes(x = x, y = y)) +
  labs(x = "TIDAL C:P", y = "Planar C:P") +
  theme_bw(base_size = 8)
)

conversion_plot + 
  geom_point(data = exp_data, aes(x = tidal, y = planar, shape = designator)) +
    scale_shape_manual(
    guide = guide_legend("", title.position = "top"),
    breaks = exp_data$designator,
    labels = c("A, B", "C, D", "E, F"),
    values = c(19, 15, 1)
  ) +
  theme(legend.position = "none")

conversion_plot +
  geom_point(data = exp_data, aes(x = tidal, y = planar, shape = designator)) +
  scale_shape_manual(
    guide = guide_legend("", title.position = "top"),
    breaks = exp_data$designator,
    labels = c("A, B", "C, D", "E, F"),
    values = c(19, 15, 1)
  ) +
  theme(legend.position = "bottom")

```