---
title: "Transform TIDAL results to planar C:P"
author: "irw"
date: "Created 20240207 -- Last compiled `r format(Sys.Date())`"
always_allow_html: false
output:
  html_document:
    df_print: paged
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
    keep_tex: false
params:
  verbose: yes
changes:
- "Adapted from report template"
editor_options: 
  chunk_output_type: inline
---

```{r, include = F}

child_chunk = knitr::opts_knit$get("child")

```


```{r, eval = !child_chunk, include=FALSE}
if (params$verbose) {
  knitr::opts_chunk$set(
    echo = TRUE,
    warning = TRUE,
    message = FALSE
  )
} else {
  knitr::opts_chunk$set(
    echo = FALSE,
    warning = FALSE,
    message = FALSE
  )
  }



knitr::opts_chunk$set(fig.width=120*1/25.4, fig.height=90*1/25.4, 
                      fig.retina = 1, dpi=500)
```


```{r, echo=FALSE}
if (knitr::is_latex_output()) {
knitr::asis_output("\\newpage")
}
```

```{r, eval = !child_chunk}
# Setup chunk

# Load libraries
library(tidyverse)
library(here)
library(jsonlite)
library(kableExtra)
library(plotly)
library(glue)
library(pracma)
library(car)

source(here::here("resources/utils/make_path.R"))
source(here::here("resources/utils/capitalize.R"))
source(here::here("resources/plot/print_ggplot.R"))
source(here::here("resources/plot/plot_interactive.R"))

source(here::here("deposition/scripts/get_summary_average_deposition.R"))
source(here::here("deposition/scripts/convert_to_planar.R"))
source(here::here("deposition/scripts/calc_cp.R"))

knitr::asis_output("# Session Info")
sessionInfo()
```

```{r, eval = !child_chunk, echo=FALSE}
if (knitr::is_latex_output()) {
knitr::asis_output("\\newpage")
}
```

```{r, child = '../deposition-summary/00-import-deposition.Rmd'}
```

```{r}
parent <- "deposition/cp-transform"
```

Original reference:

-	Schroeter JD, Pritchard JN, Hwang D, Martonen TB. Airway identification within planar gamma camera images using computer models of lung morphology. Pharm Res. 2005;22(10):1692-9.

NOTE: This is based on a symmetric airway model, using Weibel numbering.

Fraction reference:

- On the Validation of Generational Lung Deposition Computer Models Using Planar Scintigraphic Images: The Case of Mimetikos Preludium. Bo Olsson and Stavros C. Kassinos. Journal of Aerosol Medicine and Pulmonary Drug Delivery 2021 34:2, 115-123

```{r}
shallow_dates <- c(
  "20231127",
  "20231130",
  "20231204"
)
deep_dates <- c(
  "20231218",
  "20231220",
  "20231222"
)

uniform_dates <- c(
  "20240123",
  "20240126",
  "20240221"
)
```

```{r}
shallow_tidal_cp <- full_lung_aggregate %>% 
  subset(date %in% shallow_dates) %>% 
  get_summary_average_deposition(region) %>% 
  calc_cp(transform = T)

deep_tidal_cp <- full_lung_aggregate %>% 
  subset(date %in% deep_dates) %>% 
  get_summary_average_deposition(region) %>% 
  calc_cp(transform = T)

shallow_tidal_uniform_cp <- full_lung_aggregate %>% 
  subset(date %in% uniform_dates) %>% 
  get_summary_average_deposition(region) %>% 
  calc_cp(transform = T)

shallow_tidal_cp

deep_tidal_cp

shallow_tidal_uniform_cp

t.test(shallow_tidal_cp$values, deep_tidal_cp$values)

t.test(shallow_tidal_cp$values, shallow_tidal_uniform_cp$values)
```
