---
title: "Comparison to KDEP, MPPD simulations"
author: "irw"
date: "Last compiled `r format(Sys.Date())`"
always_allow_html: false
output:
  html_document:
    df_print: paged
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
    keep_tex: false
params:
  verbose: no
editor_options: 
  chunk_output_type: inline
---

```{r, include = F}

child_chunk = knitr::opts_knit$get("child")

```


```{r, eval = !child_chunk, include=FALSE}
if (params$verbose) {
  knitr::opts_chunk$set(
    echo = TRUE,
    warning = TRUE,
    message = FALSE
  )
} else {
  knitr::opts_chunk$set(
    echo = FALSE,
    warning = FALSE,
    message = FALSE
  )
}



knitr::opts_chunk$set(fig.width=120*1/25.4, fig.height=90*1/25.4, 
                      dpi=500, fig.retina = 1)
```


```{r, echo=FALSE}
if (knitr::is_latex_output()) {
knitr::asis_output("\\newpage")
}
```

```{r, eval = !child_chunk}
# Setup chunk

# Load libraries
library(tidyverse)
library(here)
library(jsonlite)
library(kableExtra)
library(plotly)
library(glue)
library(pracma)
library(car)

source(here::here("resources/utils/make_path.R"))
source(here::here("resources/plot/print_ggplot.R"))
source(here::here("resources/plot/plot_interactive.R"))

source(here::here("deposition/scripts/get_summary_average_deposition.R"))
source(here::here("deposition/scripts/calc_lr.R"))

knitr::asis_output("# Session Info")
sessionInfo()
```

```{r, eval = !child_chunk, echo=FALSE}
if (knitr::is_latex_output()) {
knitr::asis_output("\\newpage")
}
```


```{r}
parent <- "deposition/mppd-kdep"
```

```{r, child = '../deposition-summary/00-import-deposition.Rmd'}
```

```{r, child = '00-import.Rmd'}
```

```{r}
shallow_dates <- c(
  "20231127",
  "20231130",
  "20231204"
)
deep_dates <- c(
  "20231218",
  "20231220",
  "20231222"
)

uniform_dates <- c(
  "20240123",
  "20240126",
  "20240221"
)
```

## Lobe comparison (MPPD) 

MPPD lobe data includes lobar bronchus to end.

### MPPD % Deposited - Shallow

```{r}
mppd_data_shallow

knitr::asis_output("L:R ratio")

mppd_data_shallow %>% 
  mutate(
    side = case_when(
      str_detect(lobe, "l.") ~ "left",
      str_detect(lobe, "r.") ~ "right",
      .default = NA
    )
  ) %>% 
  drop_na() %>% 
  group_by(side) %>% 
  summarize(pct = sum(pct_dep),
            values = list(pct)) %>% 
  calc_lr()
```

### TIDAL % Deposited - Shallow

```{r}
full_lung_aggregate %>% 
  subset(date %in% shallow_dates) %>% 
  get_summary_average_deposition(lobe) %>% 
  mutate(
    lobe = factor(lobe, levels = c("mt", "tb", "ru", "rm", "rl", "lu", "ll"))
  ) %>% 
  arrange(lobe)

knitr::asis_output("L:R ratio")

full_lung_aggregate %>% 
  subset(date %in% shallow_dates) %>% 
  mutate(
    side = case_when(
      str_detect(lobe, "l.") ~ "left",
      str_detect(lobe, "r.") ~ "right",
      .default = "upper"
    )
  ) %>% 
  get_summary_average_deposition(side) %>% 
  calc_lr()
```

### TIDAL % Deposited - Shallow (Uniform)

```{r}
full_lung_aggregate %>% 
  subset(date %in% uniform_dates) %>% 
  get_summary_average_deposition(lobe) %>% 
  mutate(
    lobe = factor(lobe, levels = c("mt", "tb", "ru", "rm", "rl", "lu", "ll"))
  ) %>% 
  arrange(lobe)

knitr::asis_output("L:R ratio")

full_lung_aggregate %>% 
  subset(date %in% uniform_dates) %>% 
  mutate(
    side = case_when(
      str_detect(lobe, "l.") ~ "left",
      str_detect(lobe, "r.") ~ "right",
      .default = "upper"
    )
  ) %>% 
  get_summary_average_deposition(side) %>% 
  calc_lr()
```

### MPPD % Deposited - Deep

```{r}
mppd_data_deep

knitr::asis_output("L:R ratio")

mppd_data_deep %>% 
  mutate(
    side = case_when(
      str_detect(lobe, "l.") ~ "left",
      str_detect(lobe, "r.") ~ "right",
      .default = NA
    )
  ) %>% 
  drop_na() %>% 
  group_by(side) %>% 
  summarize(pct = sum(pct_dep),
            values = list(pct)) %>% 
  calc_lr()
```


### TIDAL % Deposited - Deep

```{r}
full_lung_aggregate %>% 
  subset(date %in% deep_dates) %>% 
  get_summary_average_deposition(lobe) %>% 
  mutate(
    lobe = factor(lobe, levels = c("mt", "tb", "ru", "rm", "rl", "lu", "ll"))
  ) %>% 
  arrange(lobe)

knitr::asis_output("L:R ratio")

full_lung_aggregate %>% 
  subset(date %in% deep_dates) %>% 
  mutate(
    side = case_when(
      str_detect(lobe, "l.") ~ "left",
      str_detect(lobe, "r.") ~ "right",
      .default = "upper"
    )
  ) %>% 
  get_summary_average_deposition(side) %>% 
  calc_lr()
```

## Regional comparison (KDEP)

https://journals.sagepub.com/doi/pdf/10.1177/ANIB_24_1-3  
pp. 20 (PDF)

ICRP 66 definitions:

- ET1: Nasal (anterior)
- ET2: Nasal (posterior), mouth -> larynx
- BB: Trachea (G0) - G8
- bb: G9 - G15 (respiratory bronchioles)
- AI: G16 - end

Corresponding TIDAL regions:

- MT: ET2
- TB: BB
- C: BB
- P: bb + AI

### KDEP % Deposited

```{r}
kdep_data
```

### TIDAL % Deposited - Shallow

```{r}
full_lung_aggregate %>% 
  subset(date %in% shallow_dates) %>% 
  mutate(
    icrp_region = case_when(
      lobe == "mt" ~ "et2",
      lobe == "tb" ~ "BB",
      region == "c" ~ "BB",
      region == "p" ~ "bb_AI"
    ),
    icrp_region = factor(icrp_region, levels = c("et2", "BB", "bb_AI"))
  ) %>% 
  get_summary_average_deposition(icrp_region)
```

### TIDAL % Deposited - Deep

```{r}
full_lung_aggregate %>% 
  subset(date %in% deep_dates) %>% 
  mutate(
    icrp_region = case_when(
      lobe == "mt" ~ "et2",
      lobe == "tb" ~ "BB",
      region == "c" ~ "BB",
      region == "p" ~ "bb_AI"
    ),
    icrp_region = factor(icrp_region, levels = c("et2", "BB", "bb_AI"))
  ) %>% 
  get_summary_average_deposition(icrp_region)
```

### TIDAL % Deposited - Shallow (Uniform)

```{r}
full_lung_aggregate %>% 
  subset(date %in% uniform_dates) %>% 
  mutate(
    icrp_region = case_when(
      lobe == "mt" ~ "et2",
      lobe == "tb" ~ "BB",
      region == "c" ~ "BB",
      region == "p" ~ "bb_AI"
    ),
    icrp_region = factor(icrp_region, levels = c("et2", "BB", "bb_AI"))
  ) %>% 
  get_summary_average_deposition(icrp_region)
```

